Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

Real world systems are most vulnerable to two attacks: preimage
and second preimage. For systems relying on password validation via
comparing two hashes, preimage attacks would allow inverting the hash,
increasing the risk associated with losing a database of hashed passwords.
On the other hand, systems relying on message validation, such as signature
checks or file integrity checks would be easily broken with fast second
preimage techniques. Both of these attacks are theoretically intractable, thus protecting
real-world systems. In practice, however, research efforts have been
successfully devoted to the production of collision attacks, producing
efficient results.  These attacks have real consequences for
trustworthy systems, and evaluating their impact and risk is an
important and underserved area of the existing literature.
We seek to evaluate trustworthiness by building a measure of the relative difference between a collision
attack and second preimage attacks. The literature describes a number
of successful high profile attacks against popular, and industrially
relevant hash functions. 
For MD5, these include cloning a CA certificate \cite{Stevens2009}
and creating malicious executables \cite{MD5Executables}. For SHA-1, a pair of
PDFs with the same hash was successfully produced \cite{cryptoeprint:2017:190}. Many of these
techniques, however, exploit not the hash but the flexibility of the file format
relying instead on embedding arbitrary, collidable data and later checking for
the presence of one of the matching blocks. This is especially obvious in the
MD5 executable and SHA-1 PDF attacks. The work by M. Stevens to perform the
CA Collision, on the other hand, was a chosen prefix attack extending
previous work from 2007 \cite{Stevens2007}.

@@

The following techniques have been extensively tested on MD4 and partially
tested on MD5 and are believed to apply fully to MD5. They likely apply to any
hash function with similar structure, but as of yet, are untested.

@@

We find justification for the distance heuristic, $\delta$, defined in Section
\ref{terminology:distance} in the existing literature on MD4:
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes have distance $\delta(C_{Dobbertin}, C_{Kasselman}) = 1$
under this metric. Since others have shown that MD4 is untrustworthy, we can
limit our search space to small distances, exploiting the structure of MD4. This
reduction in search space decreases the time to solve a single 3-CNF-SAT
instance at the expense of restricting ourselves to highly connected, highly
similar differential paths. Further, we can achieve parallel solving
capabilities by noting that this problem is embarrassingly parallel: each
candidate collision class can be independently checked on separate machines.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{Figs/graph-clique-crop.pdf}
\caption{}
\label{Fig:Graph-Clique}
\end{center}
\end{figure}

>>>>

Real world systems are most vulnerable to two attacks: preimage
and second preimage. For systems relying on password validation via
comparing two hashes, preimage attacks would allow inverting the hash,
increasing the risk associated with losing a database of hashed passwords.
On the other hand, systems relying on message validation, such as signature
checks or file integrity checks would be easily broken with fast second
preimage techniques. Both of these attacks are theoretically intractable, thus protecting
real-world systems. In practice, however, research efforts have been
successfully devoted to the production of collision attacks, producing
efficient results.  These attacks have real consequences for
trustworthy systems, and evaluating their impact and risk is an
important and underserved area of the existing literature.

We evaluate trustworthiness by building a measure of the relative difference between a collision
attack and second preimage attacks. The literature describes a number
of successful high profile attacks against popular, and industrially
relevant hash functions. 
For MD5, these include cloning a CA certificate \cite{Stevens2009}
and creating malicious executables \cite{MD5Executables}. For SHA-1, a pair of
PDFs with the same hash was successfully produced \cite{cryptoeprint:2017:190}. Many of these
techniques, however, exploit not the hash but the flexibility of the file format
relying instead on embedding arbitrary, collidable data and later checking for
the presence of one of the matching blocks. This is especially obvious in the
MD5 executable and SHA-1 PDF attacks. The work by M. Stevens to perform the
CA Collision, on the other hand, was a chosen prefix attack extending
previous work from 2007 \cite{Stevens2007}.

@@

We approach the problem of logical cryptanalysis by applying
techniques based in 3-CNF-SAT solving to explore the underlying
structure of various hash functions.  In this paper we present the
first case of extensive testing of these methods on the MD4 hash
function and the partial testing of these methods on the MD5 hash
function.  Our results lead us to believe that these techniques likely
fully apply to MD5, as well as any other hash function with similar structure.

@@

We justify our the distance heuristic, $\delta$, defined in Section
\ref{terminology:distance} in the existing literature on MD4.  Using
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes, we find they have distance $\delta(C_{Dobbertin}, C_{Kasselman}) = 1$
under our metric. Since others have shown that MD4 is untrustworthy,
we hypothesized (correctly) that we could
limit our search space to small distances, exploiting the structure of MD4. This
reduction in search space decreases the time to solve a single 3-CNF-SAT
instance at the expense of restricting ourselves to highly connected, highly
similar differential paths which none-the-less yield rich and
interesting sets of collisions. We further improved our performability
by exploiting the parallalizability of this problem
noting that this problem is embarrassingly parallel: allowing us to
check each
candidate collision class independently on separate machines.

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{Figs/graph-clique-crop.pdf}
\caption{Clique and pseudo-clique neighborhoods associated with
  unit distance neighborhoods of Dobbertin and Kasselman collision
  classes from the literature.}
\label{Fig:Graph-Clique}
\end{center}
\end{figure}

@@
